\subsection{Vocabulary Checklist}

\vocabularyChecklist{
    {\dtext{呀}{aa3}}
    {ss}
    {sentence suffix, to soften abruptness}
    %
    {\dtext{都}{dou1}}
    {adv}
    {also}
    %
    {\dtext{㗎}{gaa3}}
    {ss}
    {sentence suffix, fusion of \dtext{嘅}{ge3} + \dtext{呀}{aa3}}
    %
    {\dtext{敢}{gam2}}
    {sp}
    {'Well then, ...', 'Say, ...', sentence prefix resuming the thread of the previous discussion}
    %
    {\dtext{嘅}{ge3}}
    {bf}
    {noun-forming boundword; -ge added to a Verb Phrase makes it a Noun Phrase}
    %
    {\dtext{貴姓？}{gwai3sing3}}
    {ph}
    {what is (your) surname? [polite]}
    %
    {\dtext{廣東}{gwong2dung1}}
    {pw}
    {Kwangtung, a province in SE China}
    %
    {\dtext{廣東人}{gwong2dung1jan4} [\dtext{個}{go3}]}
    {n}
    {Cantonese person, person from Kwangtung Province}
    %
    {\dtext{中國人}{zung1gwok3jan4}}
    {n}
    {Chinese person}
    %
    {\dtext{佢}{keoi5}}
    {pro}
    {he, him, his}
    %
    {\dtext{佢哋}{keoi5dei6}}
    {pro}
    {they, them, their}
    %
    {\dtext{乜嘢？}{mat1je5}}
    {qw}
    {what?}
    %
    {\dtext{美國人}{mei5gwok3jan4}}
    {n}
    {American}
    %
    {\dtext{乜嘢？}{me1je5}}
    {qw}
    {what?}
    %
    {\dtext{乜嘢？}{mi1je5}}
    {qw}
    {what?}
    %
    {\dtext{呢}{ne1}}
    {ss}
    {sentence suffix for questions}
    %
    {\dtext{你}{nei5}}
    {pro}
    {you, your}
    %
    {\dtext{你哋}{nei5dei6}}
    {pro}
    {you, your (plural)}
    %
    {\dtext{我哋}{ngo5dei6}}
    {pro}
    {we, our, us}
    %
    {\dtext{朋友}{pang4jau5} \dtext{個}{go3}}
    {n}
    {friend}
    %
    {\dtext{上海}{soeng6hoi2}}
    {pw}
    {Shanghai}
    %
    {\dtext{上海人}{soeng6hoi2jan4}}
    {n}
    {person from Shanghai}
    %
    {\dtext{先生}{sin1saang1}}
    {n}
    {'Sir', term of direct address}
    %
    {\dtext{先生}{sin1saang1}}
    {n}
    {man}
    %
    {\dtext{小姐}{siu2ze2}}
    {n}
    {'Miss', Madame, term of direct address}
    %
    {\dtext{小姓}{siu2sing3}}
    {ph}
    {my surname is (polite)}
    %
    {\dtext{台山}{toi4saan1}}
    {pw}
    {Toishan, a country in southern Kwangtung about 100 miles west of Hong Kong}
    %
    {\dtext{臺山人}{toi4saan1jan4}}
    {n}
    {person from Toishan}
    %
    {\dtext{人}{jan4}}
    {n}
    {person}
    %
    {\dtext{日本人}{jat6bun2jan4}}
    {n}
    {Japanese person}
    %
    {\dtext{英國人}{jing3gwok3jan4}}
    {n}
    {Englishman, person from England}
    %
}
